### 2023-03-06 bb2.Rmd

---
title: "Old Scania Data"
author: "GB & TB"
date: '`r Sys.time()`'
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: yes
    number_sections: yes
    toc: no
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
  bookdown::word_document2:
    toc: no
    toc_depth: 2
citation_package: natbib
##classoption: titlepage
bibliography: ../Lund2205/lund.bib
titlepage: no
biblio-style: apalike
documentclass: article
fontsize: 11pt
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{graphicx}
- \usepackage{sectsty}
- \subsectionfont{\normalfont\itshape}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.height = 4)
library(eha)
source("R/tbl.R")
```

# Introduction

An interesting paper is @lygrab98.

# Read data

```{r readdata}
sca <- read.table("Data/Copy of test_tommy 2.csv", header = TRUE, sep = ";",
                  fileEncoding = "latin1")
sca$EventDate = as.Date(sca$EventDate)
sca$BirthDate = as.Date(sca$BirthDate)
sca$DeathDate <- as.Date(sca$DeathDate)
sca$LastDate <- as.Date(sca$LastDate)
sca$SexID = NULL
sca$Sex <- factor(sca$Sex)
sca$wd <- weekdays(sca$BirthDate)
sca$death_wd <- weekdays(sca$DeathDate)
sca$year <- as.numeric(format(sca$BirthDate, format = "%Y"))
n1 <- NROW(sca)
cat("No. o f rows = ", n1, "\n")
```

# Calendar time restrictions

```{r timeres}
options(warn = -1)
hog <- sca[sca$ParishName == "Hög", ]
halm <- sca[sca$ParishName == "Halmstad", ]
kag <- sca[sca$ParishName == "Kågeröd", ]
kav <- sca[sca$ParishName == "Kävlinge", ]
sir <- sca[sca$ParishName == "Sireköpinge", ]
##
hog = hog[hog$year >= 1685, ]
halm <- halm[halm$year >= 1710, ]
kag <- kag[kag$year >= 1728, ]
kav <- kav[kav$year >= 1685, ]
sir <- sir[sir$year >=  1710, ]
##
sca <- rbind(hog, halm, kag, kav, sir)
sca$LastEvent <- factor(sca$LastEvent)
xxx <- with(sca, table(wd, ParishName))
n2 <- NROW(sca)
cat("No. of rows left = ", n2, ", removed", n1-n2, "\n")
tbl(xxx, caption = "Week day by Parish.", rownames = TRUE)
options(warn = 0)
```

# Last date

```{r lastdate}
weq <- with(sca, LastEvent %in% c("Född", "Hushåll In", "Inflytt", "vigd"))
sca <- sca[!weq, ]
sca <- sca[sca$Sex != "Unknown", ]
sca$Sex <- factor(sca$Sex)
n3 <- NROW(sca)
cat("No. of rows left = ", n3, ", removed another ", n2-n3, "\n")
```

# Sex ratio

See Table \@ref(tab:sexrat).

```{r sexrat}
xx <- with(sca, table(Sex, ParishName))
rr <- xx[2, ]  /  xx[1, ]
sex.ratio <- round(rr, 2)
tbl(sex.ratio, caption = "Sex ratio by Parish", rownames = TRUE)
```

# Survival data and cohort selection

We select a cohort consisting of all born later than 31 December 1709.

```{r surv}
sca$exit <- as.numeric(with(sca, LastDate - BirthDate))
sca$enter <- 0
sca$exit[sca$exit < 0.5] <- 0.25
sca$event <- !is.na(sca$DeathDate)
sca <- age.window(sca, c(0, 365))
## Temporarily change time scale to year:           NOTE!!!!
##sca$exit <- sca$exit / 365      NOTE: KEEP DAYS AS TIME SCALE!!
sca$bdate <- toTime(sca$BirthDate)
## We do not want to cut spells at calendar time limits:
##sca <- cal.window(sca, c(1710, 1812), surv = c("enter", "exit", "event", "bdate"))
sca <- sca[sca$year >= 1710, ] # Cohort selection
sca$Parish <- factor(sca$ParishName)
n4 <- NROW(sca)
cat("No. of rows left = ", n4, ", removed another ", n3-n4, "\n")
```

# Save data

```{r savedata, echo = TRUE}
sca$priest <- sca$Parish
levels(sca$priest) <- c("HalmSir", "HogKav", "Kage", "HogKav", "HalmSir")
saveRDS(sca, file = "Rdata/sca.rds")
```

# Analysis

See Table \@ref(tab:prelanal) and Figure \@ref(fig:anfig). For crude period statistics, see Figure \@ref(fig:period).

```{r prelanal, results = 'asis'}
fit <- coxreg(Surv(exit, event) ~ Sex + I(year - 1770) + Parish, data = sca)
ltx2(fit, caption = "Cox regression, full period.", label = "tab:prelanal")
```

```{r anfig, fig.cap = "Cumulative hazard function at baseline, boys and Halmstad 1770, model based."}
plot(fit, col = "blue", xlab = "Age (years)")
```

```{r period, fig.cap = "Crude IMR by time period start."}
cuts <- c(1710, 1725, 1750, 1775, 1800, 1812)
yy <- with(sca, cbind(enter/365, exit/365, event, toTime(BirthDate))) # NOTE: /365
xx <- perstat(yy, cuts, age = c(0, 1))
plot(cuts[-length(cuts)], xx$intensity, type = "b", ylim = c(0, 0.3), xlim = c(1710, 1811),
     xlab = "Time period", ylab = "IMR", axes = FALSE, col = "blue")
axis(1, at = cuts)
axis(2, las = 1)
box()
```

# Bourgeois-Pichat and BB

Time scale days and the log-cube transform. See Figure \@ref{fig:bpscale} for a comparison. Not very linear 
in either case.

```{r bpscale, fig.cap = "Bourgeois-Pichat in red, BB in blue."}
## sca$exit <- sca$exit * 365 # ALREADY "DAY SCALE"!
post <- sca
##post <- age.window(sca, c(29, 365))
#post$enter <- post$enter - 0.01
#post$exit = post$exit - post$enter
g <- function(x) (log(1 + x))^3
post <- post[post$year >= 1765, ]
post$exit <- g(post$exit)
post$enter <- g(post$enter)
gsurv <- with(post, Surv(enter, exit, event))
rs <- risksets(gsurv, members = FALSE)
bp <- cumsum(rs$n.events) / rs$size[1]
plot(rs$risktimes, bp, type = "l", ylim = c(0, 0.30), ##xlim = c(g(29), g(365)),
     xlab = "g(Age)", ylab = "", col = "red")
bb <- cumsum(rs$n.events) / rs$size
lines(rs$risktimes, bb, col = "blue", lty = 1)
n <- length(bb)
lines(rs$risktimes[c(29, n)], bb[c(29, n)], lty = 2, col = "blue")
lines(rs$risktimes[c(29, n)], bp[c(29, n)], lty = 2, col = "red")
```

# Conditional analysis

Condition on survival to 28:th day.

```{r cond}
p2 <- age.window(post, c(g(28), g(365)))
```

And BP vs. BB.

```{r bpbbcond, fig.cap = "Bourgeois-Pichat in red, BB in blue, conditional."}
gsurv <- with(p2, Surv(enter, exit, event))
rs <- risksets(gsurv, members = FALSE)
bp <- cumsum(rs$n.events) / rs$size[1]
plot(rs$risktimes, bp, type = "l", ylim = c(0, 0.30),
     xlab = "g(Age)", ylab = "", col = "red")
bb <- cumsum(rs$n.events) / rs$size
lines(rs$risktimes, bb, col = "blue", lty = 1)
n <- length(bb)
lines(rs$risktimes[c(1, n)], bb[c(1, n)], lty = 2, col = "blue")
lines(rs$risktimes[c(1, n)], bp[c(1, n)], lty = 2, col = "red")
```

## Gompertz

```{r gompertz, fig.cap = "Gompertz fit."}
fit.g <- phreg(Surv(exit - enter, event) ~ 1, data = p2, dist = "gompertz", param = "rate")
fit.c <- coxreg(Surv(exit - enter, event) ~ 1, data = p2)
check.dist(fit.g, fit.c)
```

Good fit, see Figure \@ref(fig:gompertz)!

## Weibull

```{r weibull, fig.cap = "Weibull fit."}
fit.w <- phreg(Surv(exit - enter, event) ~ 1, data = p2)
fit.c <- coxreg(Surv(exit - enter, event) ~ 1, data = p2)
check.dist(fit.w, fit.c)
```

Bad fit, see Figure \@ref(fig:weibull).