### 2023-03-06 desc.Rmd

---
title: "Old Scania data, summary statistics"
author: "GB & TB"
date: '`r Sys.time()`'
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
  bookdown::word_document2:
    toc: no
    toc_depth: 2
citation_package: natbib
##classoption: titlepage
bibliography: ../Lund2205/lund.bib
titlepage: no
biblio-style: apalike
documentclass: article
fontsize: 11pt
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{graphicx}
- \usepackage{sectsty}
- \subsectionfont{\normalfont\itshape}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.height = 8)
library(eha)
source("R/tbl.R")
sca <- readRDS("Rdata/sca.rds")
```

# Introduction

Some statistics for Scania data and

# Cumulative hazards by parish

First the full period, Figure \@ref(fig:totcum).

```{r totcum, fig.cap = "Cumulative hazards by parish, full period."}
par(mfrow = c(3, 2))
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca[sca$Parish == "Halmstad",])
plot(fit, main = "Halmstad", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca[sca$Parish == "Hög",])
plot(fit, main = "Hög", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca[sca$Parish == "Kågeröd",])
plot(fit, main = "Kågeröd", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca[sca$Parish == "Kävlinge",])
plot(fit, main = "Kävlinge", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca[sca$Parish == "Sireköpinge",])
plot(fit, main = "Sireköpinge", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca)
plot(fit, main = "Alla", printLegend = "bottomright")
```

## Period 1710--1765

This period is depicted in Figure \@ref(fig:firstcum).

```{r firstcum, fig.cap = "Cumulative hazards by parish, 1710--1765."}
par(mfrow = c(3, 2))
sca1 <- sca[sca$year <= 1765, ]
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca1[sca1$Parish == "Halmstad",])
plot(fit, main = "Halmstad", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca1[sca1$Parish == "Hög",])
plot(fit, main = "Hög", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca1[sca1$Parish == "Kågeröd",])
plot(fit, main = "Kågeröd", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca1[sca1$Parish == "Kävlinge",])
plot(fit, main = "Kävlinge", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca1[sca1$Parish == "Sireköpinge",])
plot(fit, main = "Sireköpinge", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca1)
plot(fit, main = "Alla", printLegend = "bottomright")
```

## Period 1766--1811

This period is depicted in Figure \@ref(fig:lastcum).

```{r lastcum, fig.cap = "Cumulative hazards by parish, 1766--1811."}
par(mfrow = c(3, 2))
sca2 <- sca[sca$year > 1766, ]
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca2[sca2$Parish == "Halmstad",])
plot(fit, main = "Halmstad", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca2[sca2$Parish == "Hög",])
plot(fit, main = "Hög", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca2[sca2$Parish == "Kågeröd",])
plot(fit, main = "Kågeröd", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca2[sca2$Parish == "Kävlinge",])
plot(fit, main = "Kävlinge", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca2[sca2$Parish == "Sireköpinge",])
plot(fit, main = "Sireköpinge", printLegend = "bottomright")
fit <- coxreg(Surv(enter, exit, event) ~ strata(Sex), data = sca2)
plot(fit, main = "Alla", printLegend = "bottomright")
par(mfrow = c(1,1))
```

# Percentage deaths in the first month

```{r propd}
deaths <- sca[sca$event == 1, c("Parish", "exit")]
x <- round(prop.table(table(deaths$exit > 28)) * 100, 1)
names(x) = c("1st month", "later")
tbl(x, caption = "Percentage deaths first month, all parishes.")
```

# BP and BB

```{r bpbb, fig.cap = "Bourgeois-Pichat (BP) and Cumulative hazards (BB), full period."}
g <- function(x) {(log(1 + x))^3}
ginv <- function(x) {exp(x^(1/3)) - 1}
sca1 <- sca ## Future use!
sca$exit <- g(sca$exit) # NOTE!!
sca$enter <- g(sca$enter) # NOTE!!
sur <- with(sca, Surv(enter, exit, event))
source("../Lund2205/R/bbbp.R")
bbbp(sur)
```

## 1710--1765

```{r bpbb1, fig.cap = "Bourgeois-Pichat (BP) and Cumulative hazards (BB), born 1710--1765."}
sur <- with(sca[sca$year <= 1765, ], Surv(enter, exit, event))
bbbp(sur)
```

## 1766--1811

```{r bpbb2, fig.cap = "Bourgeois-Pichat (BP) and Cumulative hazards (BB), born 1766--1811."}
sur <- with(sca[sca$year > 1765, ], Surv(enter, exit, event))
bbbp(sur)
```

# Father's occupation (standard)

See Table \@ref(tab:occ).

```{r occ}
xx <- with(sca1, table(FatherOccupationStandard, event))
##tbl(xx, caption = "Father's occupation by survival.", rownames = TRUE)
yrk2 <- read.table("../Lund2205/yrkenklass2 utf.csv", header = TRUE, sep = ";", fileEncoding = "UTF-8")
yy <- cbind(rownames(xx), yrk2[, 2])
indx <- match(sca1$FatherOccupationStandard, yy[, 1])
sca1$socst <- yy[indx, 2]
sca1$socst <- as.factor(sca1$socst)
xy <- with(sca1, table(socst, event))

tbl(xy, caption = "socst by death, full period.", rownames = TRUE)
saveRDS(sca1, file = "Rdata/sca1.rds")
```