### 2023-03-11 Slides/slides.Rnw
first commit

\documentclass[t]{beamer}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{natbib}

%\usepackage[figuresonly]{endfloat}

%\bibliographystyle{apalike}

\usepackage{color}

\newcommand{\emp}[1]{\textcolor{blue}{#1}}

\usetheme{Singapore}% best choice.
  \setbeamercovered{transparent}%

\newcommand{\R}{{\bf R}}
\newcommand{\code}[1]{{\tt #1}}

\title{The biometric analysis of infant mortality and temperature
  \footnote{ESHD Conference, March 2022, Madrid}
}

\date{March 3, 2022}

\author{Göran Broström, CEDAR, Umeå University \\
        Tommy Bengtson, Economic History, Lund University}

<<setup, include = FALSE>>=
library(knitr)
opts_chunk$set(engine='R',dev='pdf',fig.path='figs/slide-',tidy=FALSE,fig.height=5.5,fig.width=8,strip.white=TRUE,cache=TRUE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
@

<<getgoing>>=
library(knitr)
library(eha)
library(xtable)
library(skum)
## Load remote data: NIX!
@

\begin{document}

\begin{frame}

\maketitle

\end{frame}

\begin{frame}{This is not about ...}

\begin{description}
  \item[Biometrics] is the measurement and statistical analysis of people's unique physical and behavioral characteristics.
    \item[Infant mortality]
\end{description}

\end{frame}

\begin{frame}{What it is about}

  \begin{description}
\item[Neonatal] mortality (But see Junkka et al (2021)).
\item[Postneonatal] mortality
\item[Temperature]
  \begin{itemize}
    \item \textcolor{blue}{Climate}
    \item \textcolor{blue}{Weather}
\end{itemize}
  \end{description}
  \end{frame}

\begin{frame}{Study area}

\includegraphics[width=0.9\textwidth]{myfigs/map2.png}

\end{frame}

\begin{frame}{Infant mortality data}

\emp{All births between January 1, 1895 and December 31, 1950}

\vspace{\baselineskip}

\begin{description}
 \item[year] Time varying, starting at year of birth.
 \item[sex]
 \item[exit] Number of days of follow-up (at most a year).
   \item[event] Logical, \emph{TRUE} if death is observed.
     \item[socst] Social status, five categories
       \item[parity] Birth order.
       \item[urban] Logical, \emph{TRUE} if born Skellefte{\aa} or
       Ume{\aa} small town.
\end{description}

\end{frame}

\begin{frame}{Temperature data}

From Swedish Meteorological and Hydrological Institute (\emp{SMHI}).

<<tempdata>>=
source("R/tbl.R")
tum <- read.table("~/Forskning/Data/ume_temp.csv", skip = 10, sep = ";", header = TRUE)
source("R/tbl.R")
ww <- tum[70137:(70137 + 20), ]
names(ww) <- c("Date", "Time", "Temperature", "Quality")
##tbl(ww, caption = "Raw temperature data from first week of 1923, Umeå.", fs = 10,
  ##  linesep = c("", "", "\\addlinespace"))
#print(ww[1:10, ])
@

\end{frame}

\end{document}