### 2023-03-15 Slides/slides.Rmd
more info

---
title: A hazards approach to the biometric analysis of infant mortality
author:
    - Göran Broström\inst{1}
    - Tommy Bengtsson\inst{2}
institute:
    - \inst{1}CEDAR, Umeå University (gb@ehar.se, https://github.com/goranbrostrom)
    - \inst{2}CED, Lund University (tommy.bengtsson@ekh.lu.se)
output:
  beamer_presentation:
    theme: Warsaw
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 9, warning = FALSE, comment = NA)

library(eha)
par(cex.axis = 1.2)
```

## Outline

*   Bourgeois-Pichat (1951): Biometric analysis of IM (background).
*   Hazard-based alternative.
*   Theoretical considerations
*   Real-world examples.

## The log-cube transform

\begin{columns}
\begin{column}{0.6\textwidth}
\begin{equation*}
g(t) = C \log^3(t + 1), \; 0 < t \le 365.
\end{equation*}

where $C$ is a normalizing constant:

\begin{equation*}
C = \frac{365}{\log^3(365 + 1)}
\end{equation*}
\end{column}

\begin{column}{0.5\textwidth}
\includegraphics[width=0.9\textwidth]{../paper2_files/figure-latex/logcube-1.pdf}
\end{column}

\end{columns}

## Earlier work

*   Knodel \& Kintner (1977) 
*   Wrigley (1977).
*   Lynch, Greenhouse \& Brändström (1998).
*   Bengtsson (1999).
*   Manfredini (2004).

## Study areas

\begin{columns}
\begin{column}{0.5\textwidth}
\begin{center}
Skåne 1711--1800

\includegraphics[width=0.7\textwidth]{myfigs/scania.png}
\end{center}
\end{column}
\begin{column}{0.5\textwidth}
\begin{center}
Västerbotten 1801--1950

\includegraphics[width=0.8\textwidth]{myfigs/vb.png}
\end{center}
\end{column}
\end{columns}

## Väterbotten 1861-1890

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/bioume6190-1.pdf}

## Västerbotten 1921--1950

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/bioume9120-1.pdf}

## Comparison

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/both-1.pdf}

## Cumulative hazards, neonatal mortality, Västerbotten 1861-1950

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/enof-1.pdf}

## Västerbotten, Bourgeois-Pichat's method

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/first-1.pdf}

## Västerbotten, hazards method

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/second-1.pdf}

## Västerbotten 1921--1950

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/lowumee-1.pdf}

## Exponential fits to post-neonatal data, Västerbotten

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/postneo-1.pdf}

## Scania 1711--1800

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/scahigh-1.pdf}

##

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/skum-1.pdf}

## Conclusions

*   Good