### 2023-03-29 Slides/slides.Rmd
add basics

---
title: "A hazards approach to the biometric analysis of infant mortality"
author:
- Göran Broström\inst{1}
- Tommy Bengtsson\inst{2}
institute:
- \inst{2}CED, Lund University (tommy.bengtsson@ekh.lu.se)
- \inst{1}CEDAR, Umeå University (gb@ehar.se, https://github.com/goranbrostrom)
output:
  beamer_presentation:
    theme: Warsaw
classoption: t
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 9, warning = FALSE, comment = NA)

library(eha)
par(cex.axis = 1.2)
```

## Outline

*   Bourgeois-Pichat (1951): Biometric analysis of IM (background).
*   Hazard-based alternative.
*   Theoretical considerations
*   Real-world examples.

## Bourgeois-Pichat and causes of death

*   Two categories: 
    *   Endogenous: *inherited*, *delivery*, etc.
    *   Exogenous: *accidental*, *infectious diseases*, etc.
*   How to differentiate between the two categories without infornation of causes of death?

Bourgeois-Pichat's Biometric model.

## The biometric model

Two postulates:

1.  Endogenous deaths only occur during the neonatal period (0--28 days). 
2.  On a specific time scale, exogenous infant mortality is uniformly distributed.

## The log-cube transform {.t}

\begin{columns}[T]
\begin{column}{0.6\textwidth}
\begin{equation*}
g(t) = C \log^3(t + 1), \; 0 < t \le 365.
\end{equation*}

where $C$ is a normalizing constant:

\begin{equation*}
C = \frac{365}{\log^3(365 + 1)}
\end{equation*}

\tiny
\emph{Note:} $C$ is not part of B-P's original definition.

\end{column}

\begin{column}{0.5\textwidth}
\includegraphics[width=0.9\textwidth]{../paper2_files/figure-latex/logcube-1.pdf}
\end{column}

\end{columns}

## Demonstration of the B-P plot

Assume 100 births, of which 90 survives infancy, ten death ages:

2, 6, 10, 15, 25, 41, 90, 120, 180, 319.

```{r bpex, fig.height = 4}
C <- 365 / (log(365 + 1))^3
g <- function(x) C * (log(x + 1))^3
d <- c(2, 6, 10, 15, 25, 41, 90, 120, 180, 319)
h <- seq_along(d) / 100
plot(g(d), h, type = "b", lty = 3, xlab = "Day", axes = FALSE, ylab = "Deaths / 100", 
     ylim = c(0, 0.11), xlim = c(0, 365))
axis(2, las = 1)
abline(h = 0, v = c(0, 365))
axis(1, at = g(d), labels = d)
axis(3, at = c(0, g(28), 365), labels = c(0, 28, 365))
box()
abline(v = g(28), lty = 2, col = "magenta")
text(g(12), 0.08, "Neonatal", col = "red")
text(g(130), 0.04, "Post-neonatal", col = "blue")
```

Calculation (no ties): $\frac{1}{100}, \frac{2}{100}, \frac{3}{100}, \ldots, \frac{10}{100}.$

<!--
\begin{equation*}
p_n = \frac{n}{100}, \; n = 1, \ldots 10
\end{equation*}
-->

## Demonstration of the B-P plot II

Assume 100 births, of which 90 survives infancy, ten death ages:

2, 6, 10, 15, 25, 41, 90, 120, 180, 319.

```{r bpex2, fig.height = 4}
C <- 365 / (log(365 + 1))^3
g <- function(x) C * (log(x + 1))^3
d <- c(2, 6, 10, 15, 25, 41, 90, 120, 180, 319)
h <- seq_along(d) / 100
plot(g(d), h, type = "b", lty = 3, xlab = "Day", axes = FALSE, ylab = "Deaths / 100", 
     ylim = c(0, 0.11), xlim = c(0, 365))

abline(h = 0, v = c(0, 365))
axis(1, at = g(d), labels = d)
axis(3, at = c(0, g(28), 365), labels = c(0, 28, 365))
box()
abline(v = g(28), lty = 2, col = "magenta")
text(g(12), 0.08, "Exogenous", col = "blue")
text(g(12), 0.03, "Endogenous", col = "red")
points(g(c(41, 319)), c(6/100, 10/100), col = "blue", lwd = 2)
##
fy <- function(x, xc, yc){
    yc[1] + (yc[2] - yc[1]) / (xc[2] - xc[1]) * (x - xc[1])
}
x1 <- g(41)
y1 <- 6/100
x2 <- g(319)
y2 <- 10/100
x0 <- 0
x3 <- g(319)
y0 <- fy(x0, c(x1, x2), c(y1, y2))
y3 <- fy(x3, c(x1, x2), c(y1, y2))
lines(c(x0, x3), c(y0, y3), lty = 2)
axis(2, las = 1, at = c(0, y0, y3), labels = round(c(0, y0, y3), 3))
lines(c(0, g(319)), c(y3, y3), lty = 2)
lines(c(0, 0), c(0, y0), col = "red", lwd = 3)
lines(c(0, 0), c(y0, y3), col = "blue", lwd = 3)
```

Calculation (no ties): $\frac{1}{100}, \frac{2}{100}, \frac{3}{100}, \ldots, \frac{10}{100}.$

<!--
\begin{equation*}
p_n = \frac{n}{100}, \; n = 1, \ldots 10
\end{equation*}
-->

## Explanation of the hazards plot

Assume 100 births, of which 90 survives infancy, ten death ages:

2, 6, 10, 15, 25, 41, 90, 120, 180, 319 (same as before).

```{r hazex, fig.height = 4}
C <- 365 / (log(365 + 1))^3
g <- function(x) C * (log(x + 1))^3
d <- c(2, 6, 10, 15, 25, 41, 90, 120, 180, 319)
h <- seq_along(d) / (100:91)
plot(g(d), h, type = "b", lty = 1, xlab = "Day", axes = FALSE, ylab = "Cumulative hazards", 
     ylim = c(0, 0.11), xlim = c(0, 365))
h2 <-  seq_along(d) / 100
lines(g(d), h2, type = "b", lty = 3) 
axis(2, las = 1)
abline(h = 0, v = c(0, 365))
axis(1, at = g(d), labels = d)
axis(3, at = c(0, g(28), 365), labels = c(0, 28, 365))
box()
abline(v = g(28), lty = 2, col = "magenta")
text(g(12), 0.08, "Neonatal", col = "red")
text(g(130), 0.04, "Post-neonatal", col = "blue")
ttt <- sum(g(c(41, 90, 120, 180, 319))) + 90 * (g(365) - g(28))

```

Calculation (no ties): $\frac{1}{100}, \frac{2}{99}, \frac{3}{98}, \ldots, \frac{10}{91}.$

## Explanation of the hazards plot II

Assume 100 births, of which 90 survives infancy, ten death ages:

2, 6, 10, 15, 25, 41, 90, 120, 180, 319 (same as before).

```{r hazex2, fig.height = 4}
C <- 365 / (log(365 + 1))^3
g <- function(x) C * (log(x + 1))^3
d <- c(2, 6, 10, 15, 25, 41, 90, 120, 180, 319)
h <- seq_along(d) / (100:91)
plot(g(d), h, type = "b", lty = 1, xlab = "Day", axes = FALSE, ylab = "Cumulative hazards", 
     ylim = c(0, 0.11), xlim = c(0, 365))

axis(2, las = 1)
abline(h = 0, v = c(0, 365))
axis(1, at = g(d), labels = d)
axis(3, at = c(0, g(28), 365), labels = c(0, 28, 365))
box()
abline(v = g(28), lty = 2, col = "magenta")
text(g(12), 0.08, "Neonatal", col = "red")
text(g(130), 0.04, "Exogenous", col = "blue")
ttt <- sum(g(c(41, 90, 120, 180, 319))) + 90 * (g(365) - g(28))
lambda <- 5 / ttt
x <- g(0:365)
y <- Hweibull(x, shape = 1, scale = 1 / lambda)
lines(x, y, lty = 2, col = "blue")
```

## Earlier work

*   Knodel \& Kintner (1977) 
*   Wrigley (1977).
*   Lynch, Greenhouse \& Brändström (1998).
*   Bengtsson (1999).
*   Manfredini (2004).

## Study areas

\begin{columns}
\begin{column}{0.5\textwidth}
\begin{center}
Skåne 1711--1800

\includegraphics[width=0.7\textwidth]{myfigs/scania.png}
\end{center}
\end{column}
\begin{column}{0.5\textwidth}
\begin{center}
Västerbotten 1801--1950

\includegraphics[width=0.8\textwidth]{myfigs/vb.png}
\end{center}
\end{column}
\end{columns}

## Väterbotten 1861-1890

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/bioume6190-1.pdf}

## Västerbotten 1921--1950

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/bioume9120-1.pdf}

## Comparison

Remove?

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/both-1.pdf}

## Cumulative hazards, neonatal mortality, Västerbotten 1861-1950

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/enof-1.pdf}

## Västerbotten, Bourgeois-Pichat's method

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/first-1.pdf}

## Västerbotten, hazards method

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/second-1.pdf}

## Västerbotten 1921--1950

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/lowumee-1.pdf}

## Exponential fits to post-neonatal data, Västerbotten

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/postneo-1.pdf}

## Scania 1711--1800

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/scahigh-1.pdf}

##

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/skum-1.pdf}

## Conclusions

*   Good