### 2023-04-04 Slides/slides.Rmd
updates from paper

---
title: "A hazards approach to the biometric analysis of infant mortality"
author:
- Göran Broström\inst{1}
- Tommy Bengtsson\inst{2}
institute:
- \inst{2}CED, Lund University (tommy.bengtsson@ekh.lu.se)
- \inst{1}CEDAR, Umeå University (gb@ehar.se, https://github.com/goranbrostrom)
output:
  beamer_presentation:
#      includes: 
#         header: gb_header.txt
#         before_body: gb_beforebody.txt
      theme: Warsaw
classoption: t
header-includes:
- \newcommand{\emp}[1]{\textcolor{blue}{#1}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 9, warning = FALSE, comment = NA)

library(eha)
par(cex.axis = 1.2)
```

## Outline

*   Bourgeois-Pichat (1951): Biometric analysis of IM (background).
*   Hazard-based alternative.
*   Theoretical considerations
*   Real-world examples.

## Bourgeois-Pichat and causes of death

*   Two categories: 
    *   \emp{Endogenous}: *inherited*, *delivery*, etc.
    *   \emp{Exogenous}: *accidental*, *infectious diseases*, etc.
*   How to differentiate between the two categories without infornation of causes of death?

Bourgeois-Pichat's \emp{biometric model}.

## The biometric model

Two postulates:

1.  \emp{Endogenous} deaths only occur during the \emp{neonatal} period (0--28 days). 
2.  On a specific time scale, \emp{exogenous} infant mortality is \emp{uniformly} distributed.

## The log-cube transform {.t}

\begin{columns}[T]
\begin{column}{0.6\textwidth}
\begin{equation*}
g(t) = C \log^3(t + 1), \; 0 < t \le 365.
\end{equation*}

where $C$ is a normalizing constant:

\begin{equation*}
C = \frac{365}{\log^3(365 + 1)}
\end{equation*}

\tiny
\emph{Note:} $C$ is \emp{not} part of B-P's original definition.

\end{column}

\begin{column}{0.5\textwidth}
\includegraphics[width=0.9\textwidth]{../paper2_files/figure-latex/logcube-1.pdf}
\end{column}

\end{columns}

## Demonstration of the B-P plot

Assume 100 births, of which 90 survives infancy, ten death ages:

2, 6, 10, 15, 25, 41, 90, 120, 180, 309.

```{r bpex, fig.height = 5}
par(lwd = 2)
C <- 365 / (log(365 + 1))^3
g <- function(x) C * (log(x + 1))^3
d <- c(2, 6, 10, 15, 25, 41, 90, 120, 180, 309)
h <- seq_along(d) / 100
plot(g(d), h, type = "b", lty = 1, xlab = "Day", axes = FALSE, ylab = "Deaths / 100", 
     ylim = c(0, 0.12), xlim = c(0, 365))
axis(2, las = 1)
abline(h = 0, v = c(0, 365))
axis(1, at = g(d), labels = d)
axis(3, at = c(0, g(28), 365), labels = c(0, 28, 365))
box()
abline(v = g(28), lty = 2, col = "magenta")
text(30, 0.11, "Neonatal", col = "red")
text(110, 0.11, "Post-neonatal", col = "blue")
points(g(c(41, 309)), c(6/100, 10/100), col = "red", lwd = 2, pch = 4)
```

Calculation (no ties): $\frac{1}{100}, \frac{2}{100}, \frac{3}{100}, \ldots, \frac{10}{100}.$

<!--
\begin{equation*}
p_n = \frac{n}{100}, \; n = 1, \ldots 10
\end{equation*}
-->

## Demonstration of the B-P plot II

Assume 100 births, of which 90 survives infancy, ten death ages:

2, 6, 10, 15, 25, 41, 90, 120, 180, 309.

```{r bpex2, fig.height = 5}
par(lwd = 2)
C <- 365 / (log(365 + 1))^3
g <- function(x) C * (log(x + 1))^3
d <- c(2, 6, 10, 15, 25, 41, 90, 120, 180, 309)
h <- seq_along(d) / 100
plot(g(d), h, type = "b", lty = 1, xlab = "Day", axes = FALSE, ylab = "Deaths / 100", 
     ylim = c(0, 0.11), xlim = c(0, 365))

abline(h = 0, v = c(0, 365))
axis(1, at = g(d), labels = d)
axis(3, at = c(0, g(28), 365), labels = c(0, 28, 365))
box()
abline(v = g(28), lty = 2, col = "magenta")
text(g(12), 0.08, "Exogenous", col = "blue", cex = 0.8)
text(g(12), 0.03, "Endogenous", col = "red", cex = 0.8)
points(g(c(41, 309)), c(6/100, 10/100), col = "blue", lwd = 2)
##
fy <- function(x, xc, yc){
    yc[1] + (yc[2] - yc[1]) / (xc[2] - xc[1]) * (x - xc[1])
}
x1 <- g(41)
y1 <- 6/100
x2 <- g(309)
y2 <- 10/100
x0 <- 0
x3 <- g(309)
y0 <- fy(x0, c(x1, x2), c(y1, y2))
y3 <- fy(x3, c(x1, x2), c(y1, y2))
lines(c(x0, x3), c(y0, y3), lty = 2, lwd = 1)
axis(2, las = 1, at = c(0, y0, y3), labels = round(c(0, y0, y3), 3))
lines(c(0, g(309)), c(y3, y3), lty = 2, lwd = 1)
lines(c(0, 0), c(0, y0), col = "red", lwd = 3)
lines(c(0, 0), c(y0, y3), col = "blue", lwd = 3)
```

Calculation (no ties): $\frac{1}{100}, \frac{2}{100}, \frac{3}{100}, \ldots, \frac{10}{100}.$

<!--
\begin{equation*}
p_n = \frac{n}{100}, \; n = 1, \ldots 10
\end{equation*}
-->

## The hazards plot

Assume 100 births, of which 90 survives infancy, ten death ages:

2, 6, 10, 15, 25, 41, 90, 120, 180, 309 (same as before).

```{r hazex, fig.height = 5}
par(lwd = 2)
C <- 365 / (log(365 + 1))^3
g <- function(x) C * (log(x + 1))^3
d <- c(2, 6, 10, 15, 25, 41, 90, 120, 180, 309)
h <- seq_along(d) / (100:91)
plot(g(d), h, type = "b", lty = 1, xlab = "Day", axes = FALSE, ylab = "Cumulative hazards", 
     ylim = c(0, 0.12), xlim = c(0, 365))
h2 <-  seq_along(d) / 100
lines(g(d), h2, type = "b", lty = 3, lwd = 1) 
axis(2, las = 1)
abline(h = 0, v = c(0, 365))
axis(1, at = g(d), labels = d)
axis(3, at = c(0, g(28), 365), labels = c(0, 28, 365))
box()
abline(v = g(28), lty = 2, col = "magenta")
text(g(12), 0.11, "Neonatal", col = "red")
text(g(110), 0.11, "Post-neonatal", col = "blue")
ttt <- sum(g(c(41, 90, 120, 180, 309))) + 90 * (g(365) - g(28))

```

Calculation (no ties): $\frac{1}{100}, \frac{2}{99}, \frac{3}{98}, \ldots, \frac{10}{91}.$

## The hazards plot II

<!-- Assume 100 births, of which 90 survives infancy, ten death ages:

2, 6, 10, 15, 25, 41, 90, 120, 180, 319 (same as before).
-->

```{r hazex2, fig.height = 5}
par(mai = c(1,1,1,1), lwd = 2)
C <- 365 / (log(365 + 1))^3
g <- function(x) C * (log(x + 1))^3
d <- c(2, 6, 10, 15, 25, 41, 90, 120, 180, 309)
h <- seq_along(d) / (100:91)
plot(g(d), h, type = "b", lty = 1, xlab = "Day", axes = FALSE, ylab = "Cumulative hazards", 
     ylim = c(0, 0.13), xlim = c(0, 365))

axis(2, las = 1)
abline(h = 0, v = c(0, 365))
axis(1, at = g(c(d, 365)), labels = c(d, 365))
tabb <- g(c(0, 28, 41, 90,120, 180, 309, 365))
axis(3, at = tabb, labels = round(tabb))
box()
abline(v = g(28), lty = 2, col = "magenta")
##text(g(12), 0.08, "Neonatal", col = "red")
text(g(130), 0.11, "Total")
text(g(130), 0.05, "Exogenous", col = "blue")
ttt <- sum(g(c(41, 90, 120, 180, 319) - 28)) + 90 * (g(365) - g(28))
lambda <- 5 / ttt
x <- g(0:365)
y <- Hweibull(x, shape = 1, scale = 1 / lambda)
lines(x, y, lty = 2, col = "blue")
axis(4, at = c(0, y[366], h[10]), labels = round(c(0, y[366], h[10]), 3), las = 1)
lines(c(365, 365), c(0, y[366]), col = "blue", lwd = 3)
lines(c(365, 365), c(y[366], h[10]), col = "red", lwd = 3)
text(30, 0.12, "Neonatal", col = "red", cex = 0.8)
text(110, 0.12, "Post-neonatal", col = "blue", cex = 0.8)
```

Exogenous slope = $\frac{\#\text{(post deaths)}}{\text{post exposure}} = \frac{5}{27626} \approx 0.00018$

\tiny

Post exposure = $90 (365 - 68) + (93 - 68) + (163 - 68) + \cdots + (335 - 68) \approx 27626$

## Adding the endogenous mortality

```{r hazex3, fig.height = 5}
par(mai = c(1,1,1,1), lwd = 2)
C <- 365 / (log(365 + 1))^3
g <- function(x) C * (log(x + 1))^3
d <- c(2, 6, 10, 15, 25, 41, 90, 120, 180, 309)
h <- seq_along(d) / (100:91)
plot(g(d), h, type = "b", lty = 1, xlab = "Day", axes = FALSE, ylab = "Cumulative hazards", 
     ylim = c(0, 0.13), xlim = c(0, 365))

axis(2, las = 1)
abline(h = 0, v = c(0, 365))
axis(1, at = g(c(d, 365)), labels = c(d, 365))
tabb <- g(c(0, 28, 41, 90,120, 180, 309, 365))
##axis(3, at = tabb, labels = round(tabb))
box()
abline(v = g(28), lty = 2, col = "magenta")
##text(g(12), 0.08, "Neonatal", col = "red")
text(g(130), 0.11, "Total")
text(g(130), 0.02, "Exogenous", col = "blue")
text(g(130), 0.06, "Endogenous", col = "red")
ttt <- sum(g(c(41, 90, 120, 180, 319) - 28)) + 90 * (g(365) - g(28))
lambda <- 5 / ttt
x <- g(0:365)
y <- Hweibull(x, shape = 1, scale = 1 / lambda)
lines(x, y, lty = 2, col = "blue")
axis(4, at = c(0, y[366], h[10]), labels = round(c(0, y[366], h[10]), 3), las = 1)
lines(c(365, 365), c(0, y[366]), col = "blue", lwd = 3)
lines(c(365, 365), c(y[366], h[10]), col = "red", lwd = 3)
text(30, 0.12, "Neonatal", col = "red", cex = 0.8)
text(110, 0.12, "Post-neonatal", col = "blue", cex = 0.8)
## Add endogenous:
xn <- g(d)
yn <- h - Hweibull(xn, shape = 1, scale = 1 / lambda)
lines(xn, yn, type = "l", col = "red", lty = 4)
lines(c(g(28), g(309)), c(yn[6], yn[6]), lty = 4, lwd = 1)
```

## Earlier work

*   Knodel \& Kintner (1977).
    *    Discusses effects of \emp{breastfeeding} habits.
*   Wrigley (1977).
    *    Problems with \emp{data recording}.
*   Lynch, Greenhouse \& Brändström (1998).
    *    Mentions \emp{"constant hazard"}.
    *    Suggest \emp{linear regression} methods.
*   Bengtsson (1999).
    *    Data \emp{quality}.
*   Manfredini (2004).
    *    Deals with effect of \emp{climate} variation.

## Study areas

\begin{columns}
\begin{column}{0.5\textwidth}
\begin{center}
Skåne 1711--1800

\includegraphics[width=0.7\textwidth]{myfigs/scania.png}
\end{center}
\end{column}
\begin{column}{0.5\textwidth}
\begin{center}
Västerbotten 1801--1950

\includegraphics[width=0.8\textwidth]{myfigs/vb.png}
\end{center}
\end{column}
\end{columns}

## Väterbotten 1861-1890

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/bioume6190-1.pdf}

## Västerbotten 1921--1950

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/bioume9120-1.pdf}

## Cohort effect of birth month, Västerbotten

\begin{columns}
\begin{column}{0.6\textwidth}

\includegraphics[width=0.8\textwidth,height=0.7\textheight]{../paper2_files/figure-latex/plotcohort-1.pdf}

\end{column}

\begin{column}{0.4\textwidth}

Conclusion:
\vspace{2\baselineskip}

Post-neonatal linearity is a population property, not an individual.
\end{column}

\end{columns}

## Period effect of month, Västerbotten

\begin{columns}

\begin{column}{0.6\textwidth}

\includegraphics[width=0.8\textwidth,height=0.7\textheight]{../paper2_files/figure-latex/plotcohort3-1.pdf}

\end{column}

\begin{column}{0.4\textwidth}

Looks quite proportional!

\end{column}

\end{columns}

<!--
## Comparison

Remove?

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/both-1.pdf}

## Cumulative hazards, neonatal mortality, Västerbotten 1861-1950

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/enof-1.pdf}

## Västerbotten, Bourgeois-Pichat's method

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/first-1.pdf}

## Västerbotten, hazards method

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/second-1.pdf}

## Västerbotten 1921--1950

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/lowumee-1.pdf}

## Exponential fits to post-neonatal data, Västerbotten

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/postneo-1.pdf}

-->

## Scania 1711--1800

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/scahigh-1.pdf}

## Skellefteå 1801--1820

\includegraphics[width=0.8\textwidth]{../paper2_files/figure-latex/skum-1.pdf}

## Biometric analysis on the natural time scale

Given that the exogenous mortality follows an Exponential distribution on the log-cube scale, 
it follows that the \emp{cumulative hazards} function \emp{$H$} on the \emp{natural} time scale is

\begin{equation*}
H(t; \lambda) = \lambda \log^3(t + 1), \; t \ge 0,
\end{equation*}

and it follows that the \emp{hazard} function \emp{$h$} is given by

\begin{equation*}
h(t; \lambda) = \lambda \frac{3 \log^2(t + 1)}{t + 1}, \; t \ge 0,
\end{equation*}

As $\lambda$ varies, this constitutes a \emp{proportional hazards} family of distributions.

## Exogenous mortality on the normal time scale

```{r normal, fig.height = 4}
dg <- function(x) 3 * (log(1 + x))^2 / (1 + x) * 365 / (log(366))^3
oldpar <- par(mfrow = c(1, 2), las = 1)
x <- 0:365
y <- dg(x)
plot(x, y, type = "l", xlab = "Day", ylab = "Hazards",
     axes = FALSE, col = "blue", lwd = 2, main = "Hazard function")
axis(1, at = c(0, 28, 183, 365))
axis(2, las = 1, at = 0)
box()
abline(h = 0, v = 0)
abline(v = 28, lty = 2, col = "darkgreen")
y <- g(x)
plot(x, y, type = "l", xlab = "Day", ylab  = "Cum. hazards", main = "Cumulative hazards",
     col = "blue", axes = FALSE)
axis(1, at = c(0, 28, 183, 365))
axis(2, las = 1, at = 0)
abline(h = 0, v = 0)
box()
par(oldpar)
```

## Natural vs. log-cube time scales

\emp{Västerbotten} 1861--1890.

\includegraphics[width=\textwidth]{../paper2_files/figure-latex/natlogcube2-1.pdf}

## Natural time scale and Gompertz exogenous distribution

\emp{Skåne} 1711--1800

\includegraphics[width = \textwidth]{../paper2_files/figure-latex/escu-1.pdf}

## Conclusions

*   \emp{Advantages} of the hazards method.
       *    Fits naturally into modern survival analysis, allows \emp{censored} and \emp{truncated data}, and 
       *    The \emp{exogenous} component forms a class of \emp{proportional hazards} distributions.

*   \emp{Restrictions} for both models.
    *   Models for \emp{populations}, not for individuals.
    *   \emp{Less useful} for traditional \emp{survival analysis}.
    *   The \emp{log-cube} transform is \emp{not} universal.

*   \emp{Suggestions}.
    *   Work on the \emp{natural} time scale, if access to a computer.
    *   Try \emp{standard} distributions for \emp{exogenous}  mortality. 
    *   \emp{Remember} that we are dealing with \emp{competing risks}!
        *   Requires special care.